---
title: | 
  | SME. Final work 
  |
  | Inferential Statistics.
  |
  | Sampling distributions and confidence intervals
  |
  |
author: '**Name:** ... '
date: |
  |
  | `r Sys.setlocale("LC_TIME", "C");format(Sys.Date(), "%B %d, %Y")`
  |
  | Table of contents
  |
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
    keep_tex: true
  html_document:
    highlight: espresso
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 4
---




```{r setup, include=FALSE, echo=FALSE}
# setting the options

knitr::opts_chunk$set(echo = TRUE)
options() # printing all current options
options(digits = 2) # number of digits to display

library(moments)   # needed for finding the third and fourth  moments
# Intall if not already installed

```

>     

  
\newpage

# Introduction

> In this laboratory you will have a statement and you have to solve it using R, Rmarkdown and LaTeX. 

> All the necessary formulas will we written in LaTeX. And the answers will written using the R and Rmarkdown commands required. *$(FORMULA = OUTPUT\  FROM\  CODE)$*

> You have to follow the index given and add the explanations required.

\vspace{2cm}

# The statement

> An industrial engineer has designed a machine to pack three-pound potato bags.
However, due to various reasons, such as different weights of potatoes,
filling problems, etc. is aware that the final weight of the potato bag won`t 
be exactly three kilos, but there will be random variations in this quantity.



\vspace{2cm}

## Read the  data and describe them

> 1. The data of the sample comes in the following file: *potato_bags.csv*


```{r, eval=FALSE,echo=FALSE}
# WRITE YOUR CODE HERE, DON´T FORGET TO CHANGE eval = TRUE when you want to compile the document.

# Read the data

# Number of observations

# Number of variables

DataX <- read.table(fileName,header = TRUE,col.names = c("X.sample"))

```

**QUESTIONS**

> How many data are in the fille?

> How many variables? 

>  






> 2. Find the mean, variance, standard deviation of the data in the sample

>   


```{r echo=FALSE,eval=FALSE}
# WRITE YOUR CODE HERE, DON´T FORGET TO CHANGE eval = TRUE when you want to compile the document.

```


> The mean of the sample is *$\bar{x} = FORMULA = OUTPUT\  FROM\  CODE$*,

> the variance is *$\sigma^{2} = FORMULA = OUTPUT\  FROM\  CODE$*,

> and the standard deviation is *$\sigma = FORMULA = OUTPUT\  FROM\  CODE$*.

>    

> 3. Draw the histogram of the data showing the density not the frequencies.

```{r echo=FALSE,eval=FALSE}
# WRITE YOUR CODE HERE, DON´T FORGET TO CHANGE eval = TRUE when you want to compile the document.

```


\newpage

# Sampling distribution of the Sample Mean $X_{n}$


> Find the expected value of the Sample Mean and its standard error.

>> *$E[X_{n}] = FORMULA = OUTPUT\  FROM\  CODE$*


>> and standard error *$SD(X_{n}) = FORMULA = OUTPUT\  FROM\  CODE$* 

\vspace{2cm}

> Which distribution follows the Sample Mean and why? 

**ANSWER:**  
\vspace{2cm}
>

\vspace{2cm}

> Draw the curve supposing the original population has mean $\mu = 3$ and $\sigma = 0.2$.  





```{r echo=FALSE,eval=FALSE}
# WRITE YOUR CODE HERE, DON´T FORGET TO CHANGE eval = TRUE when you want to compile the document.
```

\newpage
# Confidence interval for the mean
>1. Which distribution do you use for finding the probability of the confidence interval? Write the  formula for the scores.

**ANSWER:**  
\vspace{2cm}
$score = FORMULA$






>2. Find the corresponding quantiles using that distribution.


**ANSWER:** 
\vspace{2cm}

```{r, eval=FALSE,echo=FALSE}
# WRITE YOUR CODE HERE, DON´T FORGET TO CHANGE eval = TRUE when you want to compile the document.

```


>The value of the quantile corresponding to the significance level is $q =OUTPUT\  FROM\  CODE$



> 3. And draw it in the graphic of the distribution.


```{r, echo=FALSE,eval=FALSE}
# WRITE YOUR CODE HERE, DON´T FORGET TO CHANGE eval = TRUE when you want to compile the document.


```





\vspace{2cm}
>3. Find the margin of error and the bounds of the interval.

**ANSWER:** 
\vspace{2cm}

```{r echo=FALSE,eval=FALSE}
# WRITE YOUR CODE HERE, DON´T FORGET TO CHANGE eval = TRUE when you want to compile the document.

```


> The margin of error is $OUTPUT\  FROM\  CODE$

> The bounds of the interval for the mean: $(OUTPUT\  FROM\  CODE)$

\newpage
# Test the following hypothesis

> To check if the machine is well calibrated, take a sample of 45 bags filled with potatoes and count their weight. With this information, 

>> Does the engineer have reason to think the machine is poorly calibrated? (Please use 5% signication level).


## Design the test 


```{r echo=FALSE,eval=FALSE}
# WRITE YOUR CODE HERE, DON´T FORGET TO CHANGE eval = TRUE when you want to compile the document.


```


> Define the null hypothesis and the alternative hypothesis

> The null hypothesis is ...

> The alternative hypothesis is ...

> The significance level is ...

>   

## Find the corresponding statistic to the sample 



```{r echo=FALSE,eval=FALSE}
# WRITE YOUR CODE HERE, DON´T FORGET TO CHANGE eval = TRUE when you want to compile the document.


```


> The statistic follows the ....

> The value of the corresponding statistic is ..*$FORMULA = OUTPUT\  FROM\  CODE$*


## Find the critical value corresponding to the significance level


```{r echo=FALSE,eval=FALSE}
# WRITE YOUR CODE HERE, DON´T FORGET TO CHANGE eval = TRUE when you want to compile the document.


```

> The critical value is ..*OUTPUT\  FROM\  CODE*


## Find the p-value


```{r echo=FALSE,eval=FALSE}
# WRITE YOUR CODE HERE, DON´T FORGET TO CHANGE eval = TRUE when you want to compile the document.


```

> The p-value is ... *$FORMULA = OUTPUT\  FROM\  CODE$*


## Make the decision

> I reject/accept the null hypothesis because the ..


## Type of errors 


```{r echo=FALSE,eval=FALSE}
# WRITE YOUR CODE HERE, DON´T FORGET TO CHANGE eval = TRUE when you want to compile the document.


```



> The value of Type I error is ...*$FORMULA = OUTPUT\  FROM\  CODE$*


> The value of Type II error is ...*$FORMULA = OUTPUT\  FROM\  CODE$*

\newpage
# List of R commands

> Write here a list of the R commands you have used

# Bibliography

> In case you have used any write it here.
